Objectification in Action: Self- and Other-Objectification in Same-gender Interactions

### 2020-07-28 Final_Manuscript.Rmd
added more Zotero references, comments on intro

@robertsfredrickson Objectification Theory, suggests that women are objectified in interpersonal encounters, and that the effect might be strongest from men, but possibly also from women. Psychological researchers studying the sexual objectification of women, and the resulting self-objectification that follows, have in recent years explored the interpersonal process of the objectification - self-objectification link. For example, in a face-to-face mixed-gender dyadic study, @garcia2016objectification found that men's reported objectification of their female interaction partner during an interaction was associated with increased self-objectification reported by their partner. @gervais2020social has reviewed the literature on interpersonal self-objectification and has organized the the process in a theoretical model called the SIMO, but this model is only relevant for mixed-gender encounters. Although there is ample evidence that women objectify other women [@loughnan2015exploring; @puvia2013being], there is currently no studies investigating the process of interpersonal objectification same-gender interactions. Does objectification by a female interaction partner have the same downstream negative consequences as being objectified by a male interaction partner? For example, perhaps women could be amplifying each others' self-objectification as they get ready to go out for the night, *before* the theoretical encounters with men outlined in @gervais2020social occur.

##Self-Objectification
Self-objectification is a multidimensional process that accounts for the cognitive mechanism that translates experiences of sexualization at the cultural level [@loughnan2015exploring] to psychological (e.g., anxiety, self-esteem, authenticity, motivational states) and behavioral (cognitive performance, body monitoring) features of mental health and well-being at the individual level [@calogero2011; @moradi2008]. @calogero2011 proposes that the construct of self-objectification can be conceptualized as a learned trait. Furthermore, it can also be elicited momentarily, through the media, for example, with sexualized images in movies and magazines, which can lead to a state of self-objectification [@bernard2018bodies; @calogero2011; @moradi2008]. Being objectified by another person and possessing trait-level self-objectification (TSO) may interact to influence experiences of feeling like a body, or state self-objectification (SSO) [@fredrickson1998swimsuit, @garcia2016objectification]. Does this habit self-objectification lead women to also objectify other women, possibly to compare oneself?

##Women Objectifying Women
Some studies have found that women objectify other women to a greater extent than they objectify men [@loughnan2015exploring]. Women objectify (dehumanized) sexualized targets presented as images [@puvia2013being], but we know next to nothing about interpersonal sexual objectification among women and what the immediate effects might be. Further, do women objectify non-sexualized women they are interacting with? What might be different for same-gender interactions? @harsey2020men found that self-objectification was related to the objectification of women to a similar degree for men and women participants, although men scored high than women on a measure of belief that objectification of women is natural for men.

@@

Self-objectification has been found to occur after a mere relationship prime among women [@sanchez2008romance] because in our culture women need to look attractive to obtain and maintain successful relationship, thus, the "college relationships condition may heighten self-objectification and the evaluation of other women in a sexualized way. Researchers have also studied interpersonal objectification and self-objectification in romantic relationships [@strelan2018birds]. Evidence is mixed, and women can sometimes have positive outcomes from appearance valuation within the context of an established romantic relationship and they experience less SSO after appearance comments within romantic relationships [@meltzer2020women], less is known about first dates. But this literature is also only about heterosexual relationships. A recent study has investigated the effect of self-objectification on the reduction of relationship building skills in general (including same-sex friendships) [CITE PWQ paper I reviewed, Yoder editor]. Empirical evidence reveals that objectification manifests through inauthenticity in romantic relationships [@brunelletal2010], adverse attitudes in regard to career aspirations, and a decrease in concentration and impairment in female cognitive performance [@kahalon2018don; @quinnetal]. Both romantic relationships and relationships in general have been studies in the context of objectification, but neither of these literatures have investigated specifically same-gender relationships.

There is evidence that the experience of state self-objectification in mixed-gender contexts (stranger and romantic) and within the self in any moment (trying on a swimsuit) has negative consequences, but what about in the context of interaction with other women? Is the self-objectification experience in mixed-gender interactions the same as that experienced in same-gender interactions? The ample research demonstrating that the male gaze has a particularly detrimental effect would suggest no [@saguyetal2010, @calogero2004test; @gervais2011you; @gay2010my]. Women do objectify other women [@puvia2013being; @loughnan2015exploring; @harsey2020men], but when they do it, does it lead to self-objectification in the same way men's objectification of women does [@garcia2016objectification]? If so, does the self-objectification experienced in these same-gender interactions have the same negative consequences for authenticity in that ineraction?

##Interpersonal Objectification and Authenticity
There is some evidence that reduced authenticity is a consequence of self-objectification in the moment [@garcia2016objectification; again CITE PWQ paper I reviewed, Yoder editor]. This link has been justified by the literature on stigmatized-stigmatizer interactions, viewing the experience of being objectified in an interaction as an identity threat situation. When a woman is objectified by a man, and then experiences self-objectification, this intergroup encounter might trigger identitiy threat, but is the same negative consequence present when a woman is objectified by another woman? woman-woman interpersonal objectification processes might diverge from this process. One the one hand, she may not be having the same negative consequences the cascade from situations that threated group-based identity [@deaux1987putting; @dovidio2006nonverbal; @hebl2005promoting], but feeling like a body, rather than a full human, in an interaction may be enough to reduce feelings of authenticity and social competence, regardless of the gender of the objectifyer [@tolman2006looking; PWQ paper I reviewed, Yoder editor].

@@

###Sample Characteristics
The sample of the current study was comprised of Western women, being that sexual objectification is most prevalent in this culture [@loughnan2015], and research on objectification conducted outside of Western or Westernized countries is scarce [@moradi2008]. Because “bodies exist within social and cultural contexts, and hence are also constructed through sociocultural practices and discourses” [@robertsfredrickson, p.174], it is important to consider how diverse social identities within unique cultural contexts may inform sexual objectification phenomenon to test the cross-cultural applicability of theoretical frameworks [@loughnan2015]. Further, sexualzing experiences and self-objectification are thought to begin a very young age, and thus, researchers have only recently begun to examine such experiences among children [e.g., @holland2016; @jongenelis2014; @bury2016]. Considering the fact that the average mean age of the investigated participants of this current study was 18.85 years, research among younger and older individuals is needed, especially because self-objectification may change over time [@robertsfredrickson]. It may be valuable to question the extent to which children, adolescents, or emerging adults of different races or ethnicities are exposed to varied amounts of sexulizing content.

>>>>

@robertsfredrickson Objectification Theory, suggests that women are objectified in interpersonal encounters, and that the effect might be strongest from men, but possibly also from women. Psychological researchers studying the sexual objectification of women, and the resulting self-objectification that follows, have in recent years explored the interpersonal process of the objectification - self-objectification link (CITE INTERPERSONAL PROCESS LINKS). For example, in a face-to-face mixed-gender dyadic study, @garcia2016objectification found that men's reported objectification of their female interaction partner during an interaction was associated with increased self-objectification reported by their partner. @gervais2020social has reviewed the literature on interpersonal self-objectification and has organized the the process in a theoretical model called the SIMO, but this model is only relevant for mixed-gender encounters. Although there is ample evidence that women objectify other women [@loughnan2015exploring; @puvia2013being], there is currently no studies investigating the process of interpersonal objectification same-gender interactions. Does objectification by a female interaction partner have the same downstream negative consequences as being objectified by a male interaction partner? For example, perhaps women could be amplifying each others' self-objectification as they get ready to go out for the night, *before* the theoretical encounters with men outlined in @gervais2020social occur.

##Self-Objectification
Self-objectification is a multidimensional process that accounts for the cognitive mechanism that translates experiences of sexualization at the cultural level [@loughnan2015exploring] to psychological (e.g., anxiety, self-esteem, authenticity, motivational states) and behavioral (cognitive performance, body monitoring) features of mental health and well-being at the individual level [@calogero2011; @moradi2008]. @calogero2011 proposes that the construct of self-objectification can be conceptualized as a learned trait. Furthermore, it can also be elicited momentarily, through the media, for example, with sexualized images in movies and magazines [@morry2001magazine], which can lead to a state of self-objectification [@bernard2018bodies; @calogero2011; @moradi2008]. Being objectified by another person and possessing trait-level self-objectification (TSO) may interact to influence experiences of feeling like a body, or state self-objectification (SSO) [@fredrickson1998swimsuit, @garcia2016objectification]. Does this habit self-objectification lead women to also objectify other women, possibly to compare oneself?

##Women Objectifying Women
Past research has focused on men’s objectifying behaviors, with little attention to the role that women may also play in enacting objectification of women. @hill2008examining assessed women’s experiences of objectification from men separately from women’s experiences of objectification from women and found that women may be socialized not only to see themselves as objects but perhaps to see other women as objects as well (THEY DIDN'T DETERMINE THIS IN PERSON).
Some studies have found that women objectify other women to a greater extent than they objectify men [@loughnan2015exploring]. Women objectify (dehumanized) sexualized targets presented as images [@puvia2013being], but we know next to nothing about interpersonal sexual objectification among women and what the immediate effects might be. Further, do women objectify non-sexualized women they are interacting with? What might be different for same-gender interactions? @harsey2020men found that self-objectification was related to the objectification of women to a similar degree for men and women participants, although men scored high than women on a measure of belief that objectification of women is natural for men. ASK RANDI-WE KNOW OBJECTIFICATION HAS MORE ADVERSE CONSEQUENCES FOR WOMEN THAN MEN (Gervais, Vescio, & Allen, 2011; Moradi & Huang, 2008; Saguy, Quinn, Dovidio, & Pratto, 2010)...MENTION THAT...HOWEVER DOES WOMEN OBJECT WOMEN TO A GREATER EXTENT THAN MEN MATTER MORE?...STILL TRYING TO FIGURE THAT OUT? ADDITIONALLY @strelan2005women FOUND THAT THE MORE WOMEN SELF-OBJECTIFY, THE MORE THEY OBJECTIFY OTHER WOMEN.

@@

Self-objectification has been found to occur after a mere relationship prime among women [@sanchez2008romance] because in our culture women need to look attractive to obtain and maintain successful relationship, thus, the "college relationships condition may heighten self-objectification and the evaluation of other women in a sexualized way. Among heterosexual male and female college students, self- objectification has been found to be positively associated with the extent to which they objectify their romantic partners [@zurbriggen2011self]. Researchers have also studied interpersonal objectification and self-objectification in romantic relationships [@strelan2018birds]. Evidence is mixed, and women can sometimes have positive outcomes from appearance valuation within the context of an established romantic relationship and they experience less SSO after appearance comments within romantic relationships [@meltzer2020women], less is known about first dates. But this literature is also only about heterosexual relationships. A recent study has investigated the effect of self-objectification on the reduction of relationship building skills in general (including same-sex friendships) [CITE PWQ paper I reviewed, Yoder editor]. Empirical evidence reveals that objectification manifests through inauthenticity in romantic relationships [@brunelletal2010], adverse attitudes in regard to career aspirations, and a decrease in concentration and impairment in female cognitive performance [@kahalon2018don; @quinnetal]. Both romantic relationships and relationships in general have been studies in the context of objectification, but neither of these literatures have investigated specifically same-gender relationships.

There is evidence that the experience of state self-objectification in mixed-gender contexts (stranger and romantic) and within the self in any moment (trying on a swimsuit) has negative consequences, but what about in the context of interaction with other women? Is the self-objectification experience in mixed-gender interactions the same as that experienced in same-gender interactions? The ample research demonstrating that the male gaze has a particularly detrimental effect would suggest no [@saguyetal2010, @calogero2004test; @gervais2011you; @gay2010my]. Women do objectify other women [@puvia2013being; @loughnan2015exploring; @harsey2020men], but when they do it, does it lead to self-objectification in the same way men's objectification of women does [@garcia2016objectification]? If so, does the self-objectification experienced in these same-gender interactions have the same negative consequences for authenticity in that interaction?

##Interpersonal Objectification and Authenticity
There is some evidence that reduced authenticity is a consequence of self-objectification in the moment [@garcia2016objectification; again CITE PWQ paper I reviewed, Yoder editor]. This link has been justified by the literature on stigmatized-stigmatizer interactions (CITE LITERATURE DIDN'T COME ACROSS THIS), viewing the experience of being objectified in an interaction as an identity threat situation (MAYBE CITE MILES-MCLEAN PAPER "STOP LOOKING AT ME!"). When a woman is objectified by a man, and then experiences self-objectification, this intergroup encounter might trigger identitiy threat, but is the same negative consequence present when a woman is objectified by another woman? woman-woman interpersonal objectification processes might diverge from this process. One the one hand, she may not be having the same negative consequences the cascade from situations that threated group-based identity [@deaux1987putting; @dovidio2006nonverbal; @hebl2005promoting], but feeling like a body, rather than a full human, in an interaction may be enough to reduce feelings of authenticity and social competence, regardless of the gender of the objectifyer [@tolman2006looking; PWQ paper I reviewed, Yoder editor]. I DONT KNOW WHERE THIS PAPER IS? ASK RANDI ABOUT IDENTITY_THREAT PAPERS

@@

###Sample Characteristics
The sample of the current study was comprised of Western women, being that sexual objectification is most prevalent in this culture [@loughnan2015], and research on objectification conducted outside of Western or Westernized countries is scarce [@moradi2008], although more current work examines objectification from a cross-cultural framework [@loughnan2015, @wollast2020cultural]. Because “bodies exist within social and cultural contexts, and hence are also constructed through sociocultural practices and discourses” [@robertsfredrickson, p.174], it is important to consider how diverse social identities within unique cultural contexts may inform sexual objectification phenomenon to test the cross-cultural applicability of theoretical frameworks [@loughnan2015]. Further, sexualizing experiences and self-objectification are thought to begin a very young age, and thus, researchers have only recently begun to examine such experiences among children [e.g., @holland2016; @jongenelis2014; @bury2016]. Considering the fact that the average mean age of the investigated participants of this current study was 18.85 years, research among younger and older individuals is needed, especially because self-objectification may change over time [@robertsfredrickson]. It may be valuable to question the extent to which children, adolescents, or emerging adults of different races or ethnicities are exposed to varied amounts of sexualizing content.
